Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-11-23 paper/paper.tex
A few tweaks to paper text before generating some more figures

\subsection*{Competition}

@@

Although two of these competitive mechanisms were investigated
by \citet{simpson_simple_2011}, the first, space-based competition, offers no
mechanism by which axon growth cones are able to communicate and the second
mechanism was implemented only for a single axis in the model (EphA receptors,
expressed in a gradient along the XX retinal axis). We wondered whether a
\emph{single} signalling mechanism working alongside the `untuned'
chemoaffinity with ligand expression given by Eq.\,\ref{e:tecligexp} could
successfully reproduce the wildtype and manipulation results. To investigate,
we studied each competitive mechanism alongside chemoaffinity in turn, setting
$m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model, $m_c=m_{\!_J}=0$ to
give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a `$\mathbf{GJ}$'
model.

@@

Space based competition is a repulsive effect which acts between any two
branches, whether they originate from different retinal ganglion cells or from
the same cell. It assumes that there is some kind of self-other signal which
acts to discourage two cells from approaching each other.
%
\begin{equation}
\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W
\end{equation}
where the distance-dependent weight $W_c$ is given by
\begin{equation}
W_c(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_c}   & d_{kb} \leq 2r_c \\
     0 & d_{kb} > 2r_c
     \end{cases}
\end{equation}
for two growth cones of interaction radius $r_c$ separated by a distance
$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
branch $b$.

The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} with $m_{\!_I}=m_{\!_J}=0$ has three
free parameters; $m_{\!_G}$, $m_c$ and $r_c$. To find the best combination of
these parameters, we reasoned that evolution would optimize the arrangement of
the wildtype pattern. We performed a simulated annealing optimization to
minimize the error metric, $\epsilon$, for the wildtype pattern
(Fig.\,\ref{f:ch}E) at simulation time 1000. The parameters returned were $m_c$
= 0.09959, $m_{\!_G}$ = 0.003841 and $r_c$ =
0.39918. Fig.\,\ref{f:nonexistent_yet} shows a simulation with these parameters.

>>>>

\subsection*{Combining chemoaffinity with competition}

@@

Although two of these competitive mechanisms were investigated
by \citet{simpson_simple_2011}, the first, space-based competition, offers no
mechanism by which axon growth cones are able to communicate and the second
mechanism was implemented only for a single axis in the model (EphA receptors,
expressed in a gradient along the XX retinal axis). We wondered whether a
\emph{single} signalling mechanism working alongside the `untuned'
chemoaffinity with ligand expression given by Eq.\,\ref{e:tecligexp} could
successfully reproduce the wildtype and manipulation results. Alongside the
same chemoaffinity mechanism, we studied each competitive mechanism in turn,
setting $m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model,
$m_c=m_{\!_J}=0$ to give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a
`$\mathbf{GJ}$' model.

@@

Space based competition was implemented as a repulsive effect acting between
any two branches, whether they originated from different retinal ganglion
cells or from the same cell. An assumption is that there is some kind of
self-other signal which acts to discourage two cells from approaching each
other. A competition movement vector, $\mathbf{C}_b$, acting on branch $b$,
which is a proximity-weighted sum of the repulsion due to all nearby branches
is defined as:
%
\begin{equation}
\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W_c
\end{equation}
where $\hat{\mathbf{x}}_{kb}$ is the unit vector from branch $k$ to branch
$b$. The distance-dependent weight $W_c$ is given by
\begin{equation}
W_c(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_c}   & d_{kb} \leq 2r_c \\
     0 & d_{kb} > 2r_c
     \end{cases}
\end{equation}
for two growth cones of interaction radius $r_c$ separated by a distance
$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
branch $b$.

The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} with $m_{\!_I}=m_{\!_J}=0$ has three
free parameters; $m_{\!_G}$, $m_c$ and $r_c$. To find the best combination of
these parameters, we reasoned that evolution would optimize the arrangement of
the wildtype pattern. We performed a simulated annealing optimization to
minimize the error metric, $\epsilon$, for the wildtype pattern
(Fig.\,\ref{f:ch}E) at simulation time 1000. The parameters returned were
$m_c$ = 0.09959, $m_{\!_G}$ = 0.003841 and $r_c$ =
0.39918. Fig.\,\ref{f:nonexistent_yet} shows the $\mathbf{GC}$ simulation with
these parameters.
